### 2015-11-14 ML/hw5/hw.tex
hw5 ml

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{fancyhdr}
\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-.3in}
\setlength{\textheight}{9in}
% \pagestyle{empty}
\pagestyle{fancy}
\begin{document}

\begin{center}
{\Large Machine Learnig Homework 5} \\[.3in]
\end{center}
\lhead{Adam Kosiorek}
\rhead{IMAT: 03661883}
\vspace*{.5in}

\section*{Problem 1}

There are four names in the hat: Martin, Alexi, Igor, and Valery. The name that is randomly drawn wins the lottery. We define three different events:
\begin{description}
 \item{$X_1$} Martin or Alexi wins.
 \item {$X_2$} Alexi or Igor wins.
 \item {$X_3$} Igor or Martin wins.
\end{description}

We have $P(X_1) = P(X_2) = P(X_3) = 0.5$ and $P(X_1, X_2) = P(X_1, X_3) = P(X_2, X_3) = 0.25$, so the events are pairwise independent. But P$(X_1, X_2, X_3) = 0$, not $0.125$, as it would if the events were all mutual independent.

\section*{Problem 2}

We use the following identities:\\
\begin{equation}
Var[X] = E[X^2] + E[X]^2\label{eq:eq1}
\end{equation}
\begin{equation}
E[X+Y] = E[X]+E[Y]\label{eq:eq2}
\end{equation}
Thus we have\\
\begin{align*}
\begin{split}
Var[X+Y] &\stackrel{\eqref{eq:eq1}}{=} E\left[(X+Y)^2\right]-E[X+Y]^2\\
&\stackrel{\eqref{eq:eq2}}{=}E\left[X^2+2XY+Y^2\right] -\left(E[X]+E[Y]\right)^2\\
&\stackrel{\eqref{eq:eq2}}{=}E\left[X^2\right] + 2E[XY] + E\left[Y^2\right] -E[X]^2-2E[X]E[Y]-E[Y]^2\\
&=Var[X]+Var[Y] + 2\left(E[XY] - E[X]E[Y]\right) = Var[X]+Var[Y]+2Cov(X,Y).
\end{split}
\end{align*}

\section*{Problem 3}

\begin{equation}
Var[X + Y] = E[(X+Y)^2]-(E[X] + E[Y])^2 = E[X^2] - E[X]^2 + E[Y]^2 - E[Y]^2 + 2E[XY] - 2E[X]E[Y]
\end{equation}

And hence,

\begin{equation}
Var[X + Y ] = Var[X] + Var[Y ] + 2Cov[X,Y].
\end{equation}

\section*{Problem 4}

$Z = \begin{pmatrix} X \\ Y \end{pmatrix} \sim \mathrm{N}(\mu_Z, \Sigma_Z)$, with $\mu_Z = \begin{pmatrix} \mu_X \\ \mu_Y \end{pmatrix}$ and $\Sigma_Z = \begin{pmatrix} \sigma_X^2 & Cov(X, Y) \\ Cov(X, Y) & \sigma_Y^2 \end{pmatrix}$. Since $\rho(X, Y) = 0$ also $Cov(X, Y) = 0$ and therefore $\Sigma_Z = \begin{pmatrix} \sigma_X^2 & 0 \\ 0 & \sigma_Y^2 \end{pmatrix}$ is diagonal. We can compute the conditional probabilities $p(X|Y) \sim \mathrm{N}(\mu_{X|Y}, \Sigma_{X|Y})$ and $p(Y|X) \sim \mathrm{N}(\mu_{Y|X}, \Sigma_{Y|X})$ using the formula (4.69) from Murphy p. 111. We get:

\begin{equation}
\begin{align}
 \mu_{X|Y} = \mu_X + Cov(X, Y) \frac{1}{\sigma_y^2}(Y - \mu_Y) = \mu_X \\
 \Sigma_{X|Y} = \sigma_X^2 - Cov(X, Y) \frac{1}{\sigma_Y^2} Cov(Y, X) = \sigma_X^2
\end{align}
\end{equation}

thus $P(X|Y) = P(X)$, proves that $X$ and $Y$ are independent.

\begin{equation}
\end{equation}

\end{document}