### 2015-12-05 cv3d_seminar/presentation.tex

\documentclass[mathserif]{beamer}
\input{macro}

\mode<presentation>
{

  	\usetheme{Pittsburgh}

  	%\usetheme[numbers,
	 %pageofpages=of,% String used between the current page and the total page count.
         % bullet=circle,% Use circles instead of squares for bullets.
          %titleline=true,% Show a line below the frame title.
          %]{Singapore}

%  \setbeamertemplate{footline}[frame number]
%  \usefonttheme[onlysmall]{structurebold}
%  \setbeamercovered{dynamic}
% \setbeamercovered{transparent}

\usebackgroundtemplate{\includegraphics[height=\paperheight]{template}}

\setbeamertemplate{frametitle}{
	\vskip30pt
 	\usebeamercolor[blue!60!green]{frametitle}
  	\centering
  	\insertframetitle
	\vskip5pt
	\color{blue!60!green}{\hrule}
	}
}

\setbeamercolor{math text}{fg=orange!80!black}
\newcommand{\hl}[1]{\textcolor{red!80!black}{\underline{#1}}}

\title[short title]{SemanticPaint}
\author[short presentator]{Adam Kosiorek}
\institute[CAMP]{Computer Aided Medical Procedures \\
Technische Universit\"at M\"unchen}
\date[]{10.12.2015}

\begin{document}

% 1. introduction
%   - lots of research in registration in visual understanding since the beginning of CV
%   - none approach so far achieved robust real time segmentation
%   - this paper introduces an end-to-end pipeline for interactive registration and segmentation of 3D enviornments
%   
% 2. State of the Art
%   - 
%   
% 3. Pipeline
%   - regristration and reconstruction
%   - classification
%   - crf and mean-field inference for smoothing
%   - VOPs
% 
%  4. Results
%   - dataset for segmentation and VOPs
%   - segmentation
%   - VOPs
%   - dataset for SRF
%   - SRF
%   
% 5. Discussion
%   - applications
%   - failures
%   
% 6. Summary
%   - what it is about
%   
% 7. Future Work
%   - scaling
%   - pure geometry
%   - (verbal) priors

%---------------------------------------------------------------------------------------

\begin{frame}
\titlepage
\end{frame}

%---------------------------------------------------------------------------------------

\begin{frame}{Outline}
\tableofcontents
\end{frame}

%---------------------------------------------------------------------------------------
\section{Introduction}
\begin{frame}
\frametitle{Introduction}

\end{frame}

%---------------------------------------------------------------------------------------

\begin{frame}
\frametitle{Two columns}
\begin{columns}[t]
\begin{column}{0.45\linewidth}
Left
\end{column}
\begin{column}{0.45\linewidth}
Right
\end{column}
\end{columns}
\end{frame}

%---------------------------------------------------------------------------------------
\section{State of the Art}

\begin{frame}
\frametitle{Another section}
\end{frame}

%---------------------------------------------------------------------------------------
\begin{frame}
\frametitle{How to include an image}
\begin{center}
\includegraphics[width=0.3\linewidth]{template}
\end{center}
\end{frame}

%---------------------------------------------------------------------------------------
\end{document}