### 2015-12-06 cv3d_seminar/presentation.tex
references and outline for presentation

\documentclass[mathserif]{beamer}
\input{macro}

\mode<presentation>
{

  	\usetheme{Pittsburgh}

  	%\usetheme[numbers,
	 %pageofpages=of,% String used between the current page and the total page count.
         % bullet=circle,% Use circles instead of squares for bullets.
          %titleline=true,% Show a line below the frame title.
          %]{Singapore}

%  \setbeamertemplate{footline}[frame number]
%  \usefonttheme[onlysmall]{structurebold}
%  \setbeamercovered{dynamic}
% \setbeamercovered{transparent}

\usebackgroundtemplate{\includegraphics[height=\paperheight]{template}}

\setbeamertemplate{frametitle}{
	\vskip30pt
 	\usebeamercolor[blue!60!green]{frametitle}
  	\centering
  	\insertframetitle
	\vskip5pt
	\color{blue!60!green}{\hrule}
	}
}

\setbeamercolor{math text}{fg=orange!80!black}
\newcommand{\hl}[1]{\textcolor{red!80!black}{\underline{#1}}}

\title[short title]{SemanticPaint}
\author[short presentator]{Adam Kosiorek}
\institute[CAMP]{Computer Aided Medical Procedures \\
Technische Universit\"at M\"unchen}
\date[]{10.12.2015}

\begin{document}

% 1. introduction
%   - lots of research in registration in visual understanding since the beginning of CV
%   - none approach so far achieved robust real time segmentation
%   - this paper introduces an end-to-end pipeline for interactive registration and segmentation of 3D enviornments
%   
% 2. State of the Art
%   - 
%   
% 3. Pipeline
%   - regristration and reconstruction
%   - classification
%   - crf and mean-field inference for smoothing
%   - VOPs
% 
%  4. Results
%   - dataset for segmentation and VOPs
%   - segmentation
%   - VOPs
%   - dataset for SRF
%   - SRF
%   
% 5. Discussion
%   - applications
%   - failures
%   
% 6. Summary
%   - what it is about
%   
% 7. Future Work
%   - scaling
%   - pure geometry
%   - (verbal) priors

%---------------------------------------------------------------------------------------

\begin{frame}
\titlepage
\end{frame}

%---------------------------------------------------------------------------------------

\begin{frame}{Outline}
\tableofcontents
\end{frame}

%---------------------------------------------------------------------------------------
\section{Introduction}
\begin{frame}
\frametitle{Introduction}

\end{frame}

%---------------------------------------------------------------------------------------
\section{State of the Art}
\begin{frame}
\frametitle{Introduction}

\end{frame}

%---------------------------------------------------------------------------------------
\section{Pipeline}
\begin{frame}
\frametitle{Introduction}

\end{frame}

%---------------------------------------------------------------------------------------
\section{Results}
\begin{frame}
\frametitle{Segmentation}

\end{frame}

\begin{frame}
\frametitle{Features}

\end{frame}

\begin{frame}
\frametitle{Streaming Random Forest}

\end{frame}

%---------------------------------------------------------------------------------------
\section{Discussion and Outlook}
\begin{frame}
\frametitle{Summary}

\end{frame}

\begin{frame}
\frametitle{Failures}

\end{frame}

\begin{frame}
\frametitle{Future Work}

\end{frame}

%---------------------------------------------------------------------------------------

\begin{frame}
\frametitle{Two columns}
\begin{columns}[t]
\begin{column}{0.45\linewidth}
Left
\end{column}
\begin{column}{0.45\linewidth}
Right
\end{column}
\end{columns}
\end{frame}

%---------------------------------------------------------------------------------------
\section{State of the Art}

\begin{frame}
\frametitle{Another section}
\end{frame}

%---------------------------------------------------------------------------------------
\begin{frame}
\frametitle{How to include an image}
\begin{center}
\includegraphics[width=0.3\linewidth]{template}
\end{center}
\end{frame}

%---------------------------------------------------------------------------------------
\begin{frame}
\frametitle{References}
\fontsize{6pt}{7.2}\selectfont
\begin{itemize}
  \item Roberts, L. G. 1963. Machine perception of three-dimensional solids. Ph.D. thesis, Massachusetts Institute of Technology.

  \item Kim, B.-S. et. al. 2013. 3D scene understanding by voxel-CRF. In Proc. ICCV.

  \item Pradeep, V. et. al. 2013. Monofusion: Real-time 3D reconstruction of small scenes with a single web camera. In Proc. ISMAR.

  \item Herbst, E. et.al. 2014. Toward online 3-d object segmentation and mapping. In IEEE International Conference on Robotics and Automation (ICRA).

  \item Valentin, J. P. et. al. 2013. Mesh based semantic modelling for indoor and outdoor scenes. In Proc. CVPR.

  \item Levoy , M. et. al. 2000. The digital Michelangelo project: 3D scanning of large statues. In Proc. SIGGRAPH. ACM.

  \item Newcombe , R. A. et. al. 2011. KinectFusion: Real-time dense surface mapping and tracking. In Proc. ISMAR.

  \item Curless , B. et. al. 1996. A volumetric method for building complex models from range images. In Proceedings of the 23rd annual conference on Computer graphics and interactive techniques. ACM, 303–312.

  \item Niessner , M. et. al. 2013. Real-time 3D reconstruction at scale using voxel hashing. ACM TOG 32, 6
\end{itemize}
\end{frame}

%---------------------------------------------------------------------------------------
\begin{frame}
\frametitle{References cont'd}
\fontsize{6pt}{7.2}\selectfont
\begin{itemize}
  \item Saffari , A. et. al. 2009. On-line random forests. In IEEE ICCV Workshop.

  \item Vitter , J. S. 1985. Random sampling with a reservoir. ACM TOMS 11, 1.

  \item Lower , D. G. 1999. Object recognition from local scale-invariant features. In Proc. ICCV.

  \item Lafferty , J. et. al. 2001. Conditional random fields: Probabilistic models for segmenting and labeling sequence data.

  \item Ktahenbül, P. et. al. 2011. Efficient inference in fully connected CRFs with Gaussian edge potentials. In NIPS.

  \item Koller , D. et.al  , N. 2009. Probabilistic Graphical Models: Principles and Techniques. MIT Press

  \item Domingos, P. et. al. 2000. Mining high-speed data streams. In Proc. SIGKDD.

  \item Lai, K. et. al. 2011. A large-scale hierarchical multi-view rgb-d object dataset. In Proc. ICRA.

  \item Valentin, J. et. al. 2015. SemanticPaint: Interactive 3D Labeling and Learning at your Fingertips. SIGGRAPH.
\end{itemize}
\end{frame}

%---------------------------------------------------------------------------------------
\end{document}