### 2016-01-10 ML/hw11/hw.tex
nn homework

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{epstopdf}

\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-.3in}
\setlength{\textheight}{9in}

\pagestyle{fancy}
\begin{document}

\begin{center}
{\Large Machine Learnig Homework 11} \\[.3in]
\end{center}
\lhead{Adam Kosiorek}
\rhead{IMAT: 03661883}
\vspace*{.5in}

\section*{Problem 1}

Consider a neural network with an input layer, a hidden layer and an output layer. Assume a linear activation function in the output layer and let the output layer take form of a single neuron. The output of the network can be expressed as

\begin{equation}
 y(x) = V^T a(x) + c = V^T f(z) = V^T f(W^T x + b)
\end{equation}

where $V$ weight matrix from the hidden layer to the output, $a$ the activation of the hidden layer, $f$ the hidden layer's nonlinearity, which acts on its input element wise, $W$ the weight matrix from the input to the hidden layer, $b$ and $c$ are biases and $x$ the input vector.

Sigmoid is a scaled and translated version of tanh, since

\begin{equation}
\begin{align}
 2 \sigma(2x - 1) = \frac{2}{1+e^{-2x}} - 1 &= \frac{2e^{2x}}{e^{2x}+1} - 1 = \frac{e^{2x}-1}{e^{2x}+1} &= tanh(x)
 \end{align}
\end{equation}

therefore, if we take $W = 2 \widetilde{W}$ and $b = 2 \widetilde{b}$ it is obvious that

\begin{equation}
 2 \sigma(W^T x + b) - 1 = tanh(\widetilde{W}^T x + \widetilde{b}) 
\end{equation}

Weights $V$ and the bias $c$ of the final layer can easily account for scaling and translation of the resulting hidden's layer activation.

\section*{Problem 2}

\begin{equation}
 \begin{align}
  \frac{d}{dx} \sigma(x) = \frac{e^{-x}}{(1 + e^{-x})^2} = \frac{1 + e^{-x}}{(1 + e^{-x})^2} - \frac{1}{(1 + e^{-x})^2} = \sigma(x) - \sigma^2(x) = \sigma(x)(1 - \sigma(x))
 \end{align}
\end{equation}

\begin{equation}
 \begin{align}
  \frac{d}{dx} tanh(x) &= \frac{2 e^{2x} \left( (e^{2x} + 1) - (e^{2x} - 1) \right)}{(e^{2x} + 1)^2} \\
  &= \frac{\left( (e^{2x} + 1) + (e^{2x} - 1) \right) \left( (e^{2x} + 1) - (e^{2x} - 1) \right)}{(e^{2x} + 1)^2} \\
  &=  \frac{(e^{2x} + 1)^2 - (e^{2x} - 1)^2}{(e^{2x} + 1)^2} \\
  &= 1 - \frac{(e^{2x} - 1)^2}{(e^{2x} + 1)^2} = 1 - tanh^2(x)
 \end{align}
\end{equation}

\end{document}