### 2016-01-31 ML/hw12/hw.tex
ML hw12 - EM and GMM

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{fancyhdr}
\usepackage{graphicx}

\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\topmargin}{-.3in}
\setlength{\textheight}{9in}

\pagestyle{fancy}
\begin{document}

\begin{center}
{\Large Machine Learning Homework 12} \\[.3in]
\end{center}
\lhead{Adam Kosiorek}
\rhead{IMAT: 03661883}
\vspace*{.5in}

\section*{Problem 1}

Let $x_n, \mu_k \in \mathcal{R}^D$ and $\Sigma_k = \Sigma = \sigma^2I \in \mathcal{R}^{D \times D}$. Since $\Sigma$ is fixed we do not reestimate it. The probability of observing a single samples $x_n$ simplifies to 
\begin{equation}
 p(x_n|\mu_k, \Sigma) = \frac{1}{(2 \pi )^\frac{D}{2} \sigma} e^{-\frac{1}{2 \sigma^2} ||x_n - \mu_k||^2}
\end{equation}

It follows that the responsibilities $r_{nk}$ are given by a softmax function, that is

\begin{equation}
 r_{nk} = \frac{ \pi_k e^{-\frac{1}{2 \sigma^2} ||x_n - \mu_k||^2} } { \sum_j \pi_j e^{-\frac{1}{2 \sigma^2} ||x_n - \mu_j||^2}} = 
 \begin{cases}
  1 & \text{if } k = \arg \min_j ||x_n - \mu_j|| \\                                                                                                                                  
  0 & \text{otherwise}                                                                                                                                 \end{cases}
\end{equation}

which results in hard assignments of samples to clusters. According to (9.40) in Bishop, the expected value of the complete-data log likelihood is given by

\begin{equation}
 \begin{align}
  L = \sum_{n=1}^{N} \sum_{k=1}^{K} r_{nk} \left( \log{\pi_k} + \log{\mathcal{N}(x_n|\mu_k, \Sigma_k)} \right)
 \end{align}
\end{equation}

In the limit, the Gaussian distribution simplifies to

\begin{equation}
 \lim_{\sigma \rightarrow 0} \mathcal{N}(x_n|\mu_k, \Sigma_k) = \lim_{\sigma \rightarrow 0} -\log{\sigma} - \frac{1}{2 \sigma^2} ||x_n - \mu_k||^2 + \text{const.} \propto -||x_n - \mu_k||^2
\end{equation}

Finally, maximizing the the log-likelihood, given by the formula below, corresponds to the minimization of the K-Means error function.

\begin{equation}
 \begin{align}
  L = \sum_{n=1}^{N} \sum_{k=1}^{K} -r_{nk} ||x_n - \mu_k||^2
 \end{align}
\end{equation}

\section*{Problem 2}

\begin{equation}
 p(x) = \sum_k \pi_k \mathcal{N}(x|\mu_k, \Sigma_k)
\end{equation}

Suppose we have two random variables of the same dimensionality: $X \sim \mathcal{N}(x|\mu_x, \Sigma_x)$ and $Y \sim \mathcal{N}(y|\mu_y, \Sigma_y)$. Let $z = \begin{pmatrix} x \\ y \end{pmatrix} \sim \mathcal{N}(x|\mu_z, \Sigma_z)$ with $\mu_z = \begin{pmatrix} \mu_x \\ \mu_y \end{pmatrix}$ and $\Sigma_z = \begin{pmatrix} \Sigma_x & 0 \\ 0 & \Sigma_y \end{pmatrix}$ We can construct a matrix $A$ such that $Az = \pi_x x + \pi_y y$ It follows that $Az \sim \mathcal{N}(\mu_{xy}, \Sigma_{xy})$ with $\mu_{xy} = \pi_x \mu_x + \pi_y \mu_y$ and  $\Sigma_{xy} = \pi_x^2 \Sigma_x + \pi_y^2 \Sigma_y$. This scheme generalizes to an arbitrary number of components. Specifically, a Gaussian Mixture Model with $K$ components is characterized by the following values:

\begin{equation}
  \begin{align}	
    E[x] &= \sum_{k=1}^K \pi_k \mu_k \\
    Cov(x) &= \sum_{k=1}^K \pi_k^2 \Sigma_k
  \end{align}
\end{equation}

\end{document}